Linear analysis of the vertical shear instability in astrophysical disks

### 2014-12-30 linear_adia.tex
added analytical derivation of critical beta critical (initial order of mag est may not be quite right, but no contradictions)

\subsubsection{Qualitative condition for buoyancy to be negligible in
  the presence of thermal relaxation}
We have shown that when $\gamma>1$ so that there is a positive
vertical entropy gradient and perturbations evolve
adiabatically ($t_c\to\infty$) that the VSI is strongly
surpressed. This is the limit where the term proportional to
$t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
becomes negligible compared to the term proportional to
$\gamma/\Gamma-1$. The latter term is associated with the vertical
entropy gradient, or buoyancy frequency squared, $N_z^2$.

@@

>>>>

\subsubsection{Qualitative condition for buoyancy to be negligible in
  the presence of thermal relaxation}
{\bf this section might be wrong, but keep for reference for now}
We have shown that when $\gamma>1$ so that there is a positive
vertical entropy gradient and perturbations evolve
adiabatically ($t_c\to\infty$) that the VSI is strongly
surpressed. This is the limit where the term proportional to
$t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
becomes negligible compared to the term proportional to
$\gamma/\Gamma-1$. The latter term is associated with the vertical
entropy gradient, or buoyancy frequency squared, $N_z^2$.

@@

\subsubsection{Upper limit on the thermal relaxation timescale for the
  fundamental isothermal VSI mode}
Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
fundamental isothermal VSI tends to a constant for large
$\khat$. Using this clue, we can calculate this critical $\beta$ by
considering $\khat\to\infty$. We proceed as follows.

At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
is real. The dispersion relation for the fundamental mode is then
\begin{align}\label{relax_disp_fund}
  0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
\end{align} 
Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} in
the limit $\khat\to\infty$ we find
\begin{align}
  0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
  \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
  &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
  0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3.\label{relax_cond2}
\end{align}
We seek solutions with
\begin{align}
  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \khat\hat{\omega} \text{ finite}.
\end{align}
Then
\begin{align}
  \khat\hat{\omega} = \frac{(\gamma-1)\beta}{\epsilon q}
\end{align}
from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
that
\begin{align}
  \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2}. 
\end{align}
For a thin disk $\epsilon\ll 1$, we infer 
\begin{align}
  \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
  \beta_\mathrm{crit} 
\end{align}
is required for the fundamental isothermal VSI to operate.  For
$\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$ we find
$\beta_\mathrm{crit}=0.125$, in agreement with Fig. \ref{relax_bcrit}
at large $\khat$.
